Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-04-02 manuscript/manuscript.Rmd
pass for nitpicks and typos

Imagine you are a graduate student, and you run a replication of a study that you are interested in building upon in your research.
The replication fails. Perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely overlaps 0, with a $p$-value of .3.
Or perhaps the interaction is numerically in the wrong direction and the main effects look different.
<!--Or maybe the main effect of interest may be null, but you found a significant effect interaction that wasn't hypothesized.-->
Whatever the details of the replication failure, you are left with a question: Should you try again and run a re-replication, or should you give up and pick a different study to extend?

@@

Replications can fail for many reasons, having to do with the original study or the replication. 
<!-- In principle, any difference between the original study and replication could be a root cause of the difference in their outcomes. But in practice, the relationship between original and replication is not symmetric, and many causes of replication failure can be attributed either more to the original study or more to the replication.  -->
Starting with the original study, one potential reason for replication failure is that the original result is fragile in some way. It could be a statistically unlikely result that was achieved through chance or $p$-hacking, or it could be sensitive to the exact conditions and time it was run under ("hidden moderators"). Another possibility is that the original reported effect size might be inflated due to some combination of heterogeneity, $p$-hacking, low power, and publication bias. A consequence of inflated effect sizes is that replication studies might select a sample size with appropriate sttaistical power for the reported effect size and then have too small a sample to reliably detect the (true) smaller effect. Replication studies might also end up underpowered due to unexpected attrition or noise in the replication sample.

Other reasons for replication failure are also primarily due to the replication study. 
A replication could observe different results from the original due to methodological differences. 
These differences in design could be intentional adaptations, unavoidable changes from lack of access to the original instructions or materials, or changes that were not expected to be critical. 
A difference in the details of recruitment (e.g., online versus in-person) might mean that particular details of a study implementation would no longer be appropriate, or the data quality controls and attention checks might be inadequate. 
If there are differences in time, place, or subject population between the two studies, corresponding changes to the materials, instructions, or procedure could be needed to provide an adaptation of the experimental paradigm to a new context. 
Such adaptation creates yet more challenges in interpretation: too many changes could cause differing results, but so could not enough changes. 
In sum, when replications fail, we generally do not know why they failed, although we may speculate.

@@

<!-- Re-replication has not been studied much, so it is uncertain how often re-replications recover the results of an original study. That is, it is not well known how definitive a first replication failure is.  -->
Relatively few studies have estimated re-replication success. 
One estimate comes from @ebersole2020. After the original RP:P result of a replication rate around 40% [@openscienceconsortium], @gilbert2016 raised concerns that methodological differences might explain low replication rate. In 11 of the RP:P studies, the original authors had raised concerns about protocol fidelity prior to data-collection.
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication. 
Thus, @ebersole2020 re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts. This study thus tested whether larger sample sizes and/or "better" methodologies would "rescue" the failed RP:P replications.
The result was that 2 of the 10 revised protocols found a significant result (but not the 1/11 that originally was successful).
Even when a significant effect was recovered, the effect size was much diminished [@ebersole2020].
<!-- We could frame @ebersole2020 as an attempt to recover the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks. -->
The re-replication success rate was thus 2/9 (22%) for studies with non-significant first replications, under favorable, high-resource circumstances.

@@

While @ebersole2020 looked at re-replication success rate in a high resource setting with access to large, multi-site samples and expert input, most early career scientists do not have those kinds of resources to devote to re-replication attempts. On the other hand, early career researchers may also be prone to a different distribution of causes of initial replication failure. Thus, it is an open question what the re-replication rate is for failed replications of early career researchers. Further, estimating this rate could inform researhcers' decisions on what to do next after failed replications.

Here, we investigate re-replication in the context of graduate student replication class projects.
Students in a graduate methods class on experimental methods in psychology were given the option to "rescue" an experiment that a student in a prior year had previously failed to replicate as their class project [first replications reported in @boyce2023].
The focus of the class was best practices in experimental design; thus, students looked at both the original paper and the failed replication to examine potential issues in the replication that they could fix in their rescue project. 
In the present paper, we report the results of `r num_proj` re-replication rescue projects. We find that a minority of projects (5/17, 29%) successful recovered the original effect.

@@

PSYCH 251 is a graduate-level experimental methods class taught by MCF.
In previous years, students have conducted replication projects [as reported in @boyce2023 and @hawkins2018].
In Fall 2023, students were offered the option to complete a "rescue" project where they re-replicated one of the unsuccessful replications from a previous year.
Students could also opt to do a normal replication instead.
If they opted for a rescue, they were given the option to participate in a collaborative meta-science project and earn authorship on the current paper based on 1) completing a rescue project, 2) reviewing another student's rescue project for analytic errors, and 3) reviewing and approving the final manuscript. 
We report on the result of `r num_proj` rescue projects that opted to be part of the paper and completed data collection.

A spreadsheet of projects, individual project write-ups (both first replications and rescues), links to individual project data and analyses for rescue projects, and the analytic code for this paper are all available at https://osf.io/cyk5w/.

@@

The experiments that were re-replicated were chosen from studies that failed to replicate in @boyce2023.
We created an initial list of `r nrow(filtered)` rescue-eligible studies that 1) had received a subjective replication success score of 0, .25 or .5 (on a 0--1 scale) in @boyce2023, 2) had a Github repository available (github repositories were used in class starting in academic year 2015-2016), and 3) where the original experiment had 200 or fewer participants (to ensure we could afford to match or increase the sample size based on the class budget).
We then contacted the replication project authors for permission to share their report and repository with a new student and include it as a supplement on a resulting paper.
This left `r length(approved)` options for the students to choose from.
`r length(chosen)` students chose to do rescue projects, and
`r length(chosen)-num_proj` students did not finish their project or did not indicate interest in being part of this paper, leaving a final sample of `r num_proj` rescue projects.

@@

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions TODO LINK was given to students.
In general, we encouraged students to add manipulation checks as appropriate, and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of @paxton2012 switched from using the cognitive reflection task (CRT) which many online participants are over-familiar with for the newer CRT-2 which is less overused but attempts to measure the same construct [@thomson2016].
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
@tarampi2016 had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map. The replication ported this experiment online and had participants click a drop down menu to select left or write for each turn. The rescue removed the unwieldy dropdown menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

@@

We followed @boyce2023 in the properties of the studies we measure and how we quantify replication success. 
Each project was rated on the basis of subjective replication success (on a 0--1 scale) by both MCF and one of VB and BP.
We thought about replication in terms of general match in key study results (direction, magnitude, and significance), rather than focusing on any singular numeric result or significance cut-off.
Interrater reliability was `r irr |> round(3)`; disagreements were minor (at no point greater than .25) and were resolved through discussion.

@@

We measure statistical consistency using $p$-original, the $p$-value on the null hypothesis that one study's population effect comes from the same distribution of population effects as another study or studies [@mathurNewStatisticalMetrics2020a] If we assume no heterogeneity between studies, then this is the same as asking how consistent one study's estimate of the population-level effect is with another study's estimate of the population-level effect.
However, meta-analyses of closely related studies, such as those from different sites of multi-site replications, indicate that there is often heterogeneity between studies.
$p$-original allows for such heterogeneity, as governed by a parameter $\tau$.
We cannot accurately estimate the level of heterogeneity for an effect meta-analytically as we have only a small number of replications per study.
Instead, we impute a heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson-collentine2020 in prior multi-site replications in psychology.

@@

For the original study and each replication, we recorded the number of participants post-exclusions.
For studies where some extra conditions were dropped in some replications, we counted only the participants in the key conditions all replications had for comparability.
For instance, if an original study compared between two critical conditions but also had a baseline control, we did not count the participants in the baseline condition if a replication did not include this condition.
We also recorded whether each study was conducted online using crowd-sourced platform or not.

@@

Our primary question of interest was how many of the `r num_proj` rescue projects succeeded at replicating the results in the original study.
Across the `r nrow(subj)` rescue projects, `r nrow(success)` mostly or fully replicated the original results according to the subjective replication ratings.
`r pluck(tallied,2,1)` had a rating of 0, `r pluck(tallied, 2,2)` got a rating of .75, and `r pluck(tallied, 2,3)` got a rating of 1.
Thus, a first pass answer to the question "how often can a failed replication be salvaged?" is `r round(nrow(success)/nrow(subj)*100)`% (bootstrapped 95% CI: `r round(bootstrap[2]*100)`% - `r round(bootstrap[3]*100)`%) of the time.

@@

smd_scale_small <- ggplot(for_plotting_d |> anti_join(big),aes(x=reorder(name_pretty, point, FUN=max),y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
  scale_shape_manual(values=shape)+
  theme(legend.position = "bottom", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standarized mean difference", x="")

level_order = c("Yeshurun & Levy 2003",
                "Ngo et al. 2019", 
                 "Krauss & Wang 2003",
                 "Jara-Ettinger et al 2022")
smd_scale_big <- ggplot(for_plotting_d |> inner_join(big) |> 
                          mutate(name_pretty=factor(name_pretty, levels=level_order)), aes(x=name_pretty,y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
    scale_shape_manual(values=shape)+
  theme(legend.position = "none", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standarized mean difference", x="")

@@

As a complement to subjective replication ratings, we also statistically compared the effect sizes of the rescue, first replication, and original study on one preregistered key measure per study [again following @boyce2023].
Though we did not conduct a systematic review of the literature to identify all possible replications, we also consider the effect sizes obtained in additional replications when we were aware of them (either from other class projects, or external replications in the literature),

@@

all |> ungroup() |>  left_join(names) |> select(Paper=name_pretty, `All reps`=orig_v_other, "Rescue" = orig_v_rescue, "Non-rescue"=orig_v_not_rescue, "Other reps"=rescue_v_reps) |> knitr::kable(format="latex", booktabs=T, linesep="", caption="P-original values between different sets of experiments. The primary analysis is between the original result and the meta-analytic aggregation of all replications. All p-originals assume an imputed heterogeneity value of tau=.21.", digits=3) |> 
  add_header_above(c(" "=1, "Original and"=3, "Rescue and"=1)) |> add_header_above(c(" "=1, "P-original comparing between"=4))

@@

Scientists' intuitions about whether a replication is successful and whether an effect provides support for a hypothesis (including heuristic cutoffs like p\<.05) do not always align with measures of statistical consistency [@patil2016].
For instance, two studies may both find that condition 1 results in a significantly higher outcome measure than condition 2, but the effect magnitudes may be sufficiently different that they are statistically unlikely to have come from the same population.
On the other hand, one study may find a statistically significant, but imprecisely estimated effect in a small sample, and second study may find a near-zero (null) effect, but the effect estimates from the two studies may be statistically compatible, despite one supporting a hypothesized difference and the other not.

As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found).
The value of $p$-original represents how likely an effect size equal to or more extreme than the original effect size is to occur if it comes from a distribution of population effect sizes defined by the mean and standard error of the replications (combined meta-analytically) with a between-population standard deviation of $\tau=.21$ [imputed from @olsson-collentine2020].
That is, it $p$-original is a measure of the statistical consistency of the original result with the totality of the replications, assuming a given level of heterogeneity of effects between studies.
Smaller $p$-original values indicate more inconsistency.

@@

As secondary measures, we calculated the $p$-original values between a) the original and the rescue, b) the original and non-rescue replications, and c) the rescue and other replications.
For a) the original versus the rescue, median value of $p$-original was `r rescue_summ$median |> round(2)` [IQR: `r rescue_summ$low |> round(2)` - `r rescue_summ$high|> round(2)`], and `r round(rescue_summ$less_than*100)`% of the $p$-original values were less than .05.
For b) the original versus all the non-rescue replications, median value of $p$-original was `r rep_summ$median |> round(2)` [IQR: `r rep_summ$low |> round(2)` - `r rep_summ$high |> round(2)`], and `r round(rep_summ$less_than*100)`% of the $p$-original values were less than .05.
For c) the rescue versus the other replications, median value of $p$-original was `r rescue_rep_summ$median |> round(2)` [IQR: `r rescue_rep_summ$low |> round(2)` - `r rescue_rep_summ$high |> round(2)`], and `r round(rescue_rep_summ$less_than*100)`% of the $p$-original values were less than .05.

@@

Overall, allowing for a heterogeneity level of $\tau$=.21 SMD, a number of original effects are not statistically consistent with rescues and replication effects.
The pattern of inconsistency does not align with which studies were rated as having replicated.
In all cases, the point estimate of the re-replication is smaller than, or in the opposite direction of, the original effect on the key measure of interest.

@@

Given the mix of successful and unsuccessful rescues, we discuss a few projects where we have speculations about why they turned out the way they did.

The rescue of @krauss2003 was successful despite an unsuccessful replication.
This study looked at the influence of a guided thinking on whether or not people gave correct justifications (drawn or written) for their answer on the Monty Hall problem.
The original paper reported correct justification from 2/67 (`r round(2/67*100)`%) in the control condition and 13/34 (`r round(13/34*100)`%) in the guided thinking condition.
The first replication struggled to recruit participants who were naive to the problem (an exclusion criterion), and many participants give very short text responses in the provided text box (only textual responses were allowed).
The replication found 0/8 correct justifications in the control and 0/11 in the guided thinking condition.
While we can't know for sure what caused the non-replication, there were clear problems observable from the small final sample and low-quality responses.
The rescue targeted these issues by adding a pre-screening that ensured participants were not familiar with the Monty Hall problem, switched the name of the problem (to reduce googling for answers), and had participants upload drawings for their justifications.
Collectively, we believe these changes brought the rescue closer to the intent of the original.
The rescue had 1/40 (`r round(1/40*100)`%) correct justifications in the control group and 6/35 (`r round(6/35*100)`%) in the guided thinking group.
This effect was significant, though smaller than the effect in the original.

@@

\@child2018 studied reading times for passages of text with different emotional valences written from either a third-person or second-person perspective.
They found a significant interaction between perspective and valence: participants read positively-valenced passages faster than negatively-valenced passages, but only when they were written from a second-person perspective.
The first replication used a similar sample size to the original study (40) and found an effect that trended in the same direction, but was substantially smaller.
The rescuer identified several possible reasons for this failure, foremost among them that the replication was conducted online while the original was conducted in person.
It is reasonable to expect online studies of reading time to be noisier than in-person studies because the environments participants completed the study in varied much more online.
The rescue attempt was still online, but used a much larger sample size to account for noisier data.
Still, the rescue found a non-significant effect that trended in the opposite direction from the original and first replication.
Even though the first replication trended in the same direction and there was an identifiable reason for the smaller effect size, increasing sample size did not rescue the effect.

@@

We opened  with a question about what an early-career psychology researcher should do given a failed replication: should they try again or give up and move on?
In our sample, the odds of a re-replication working are low [consistent with @ebersole2020].
Especially if there is not a clear, identifiable candidate for replication failure, it may be more efficient to select another result to build on.

@@

We thank the students from previous years of PSYCH 251 who graciously agreed to let us re-replicate their projects and shared their work with us. We thank the other members of 251 teaching team and students for feedback on projects. Data collection was supported by funding from the Stanford Departments of Psychology and Philosophy, Graduate School of Education, and the Symbolic Systems program.

@@

VB and MCF contributed to conceptualization. BP contributed to project administration, VB contributed to formal analysis, and MCF contributed to supervision.

Authors other than VB, BP, and MCF each designed, conducted, analyzed, and wrote up one of the projects -- these authors contributed to investigation, formal analysis, and visualization.

VB wrote the first draft, BP and MCF did editing, and all authors reviewed the final manuscript.

>>>>

Imagine you are a graduate student, and you run a replication of a study that you are interested in building upon in your research.
The replication fails. Perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely overlaps 0, with a $p$-value of .3.
Or perhaps the interaction is numerically in the wrong direction and the main effects look different.
<!--Or maybe the main effect of interest may be null, but you found a significant effect interaction that wasn't hypothesized.-->
Whatever the details of the replication failure, you are left with a question: Should you try again and run a re-replication, or should you give up and pick a different study to build upon?

@@

Replications can fail for many reasons, having to do with the original study or the replication. 
<!-- In principle, any difference between the original study and replication could be a root cause of the difference in their outcomes. But in practice, the relationship between original and replication is not symmetric, and many causes of replication failure can be attributed either more to the original study or more to the replication.  -->
Starting with the original study, one potential reason for replication failure is that the original result is fragile in some way. It could be a statistically unlikely result that was achieved through chance or $p$-hacking, or it could be sensitive to the exact conditions and time it was run under ("hidden moderators"). Another possibility is that the original reported effect size might be inflated due to some combination of heterogeneity, $p$-hacking, low power, and publication bias. A consequence of inflated effect sizes is that replication studies might select a sample size with appropriate statistical power for the reported effect size and then have too small a sample to reliably detect the (true) smaller effect. Replication studies might also end up underpowered due to unexpected attrition or noise in the replication sample.

Some other reasons for replication failure are  primarily due to the replication study. 
A replication could observe different results from the original due to methodological differences. 
These differences in design could be intentional adaptations, unavoidable changes from lack of access to the original instructions or materials, or changes that were not expected to be critical. 
A difference in the details of recruitment (e.g., online versus in-person) might mean that particular details of a study implementation would no longer be appropriate, or the data quality controls and attention checks might be inadequate. 
If there are differences in time, place, or subject population between the two studies, corresponding changes to the materials, instructions, or procedure could be needed to provide an adaptation of the experimental paradigm to the new context. 
Such adaptation creates yet more challenges in interpretation: too many changes could cause differing results, but so could not enough changes. 
In sum, when replications fail, we generally do not know why they failed, although we may speculate.

@@

<!-- Re-replication has not been studied much, so it is uncertain how often re-replications recover the results of an original study. That is, it is not well known how definitive a first replication failure is.  -->
Relatively few studies have estimated re-replication success. 
One estimate comes from @ebersole2020. After the original RP:P study reported of a replication rate around 40% [@openscienceconsortium], @gilbert2016 raised concerns that methodological differences might explain the low replication rate. In 11 of the RP:P studies, the original authors had raised concerns about protocol fidelity prior to data collection.
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication. 
Thus, @ebersole2020 re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts. This study thus tested whether larger sample sizes and/or "better" methodologies would "rescue" the failed RP:P replications.
The result was that 2 of the 10 revised protocols found a significant result (but not the 1/11 that originally was successful).
Even when a significant effect was recovered, the effect size was much diminished [@ebersole2020].
<!-- We could frame @ebersole2020 as an attempt to recover the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks. -->
The re-replication success rate was thus 2/9 (22%) for studies with non-significant first replications, under favorable, high-resource circumstances.

@@

While @ebersole2020 looked at re-replication success rate in a high resource setting with access to large, multi-site samples and expert input, most early career scientists do not have those kinds of resources to devote to re-replication attempts. On the other hand, early career researchers may also be prone to a different distribution of causes of initial replication failure. Thus, it is an open question what the re-replication rate is for failed replications by early career researchers. Further, estimating this rate could inform researchers' decisions on what to do next after failed replications.

Here, we investigate re-replication in the context of graduate student replication class projects.
Students in a graduate-level class on experimental methods in psychology were given the option to "rescue" an experiment that a student in a prior year had previously failed to replicate as their class project [first replications reported in @boyce2023].
The focus of the class was best practices in experimental design; thus, students looked at both the original paper and the failed replication to examine potential issues in the replication that they could fix in their rescue project. 
In the present paper, we report the results of `r num_proj` re-replication rescue projects. We find that a minority of projects (5/17, 29%) successfully recovered the original effect.

@@

PSYCH 251 is a graduate-level experimental methods class taught by MCF.
In previous years, students have conducted replication projects [@boyce2023; @hawkins2018].
In Fall 2023, students were offered the option to complete a "rescue" project where they re-replicated one of the unsuccessful replications from a previous year.
Students could also opt to do a normal replication instead.
If they opted for a rescue, they were given the option to participate in a collaborative meta-science project and earn authorship on the current paper based on 1) completing a rescue project, 2) reviewing another student's rescue project for reproducibility, and 3) reviewing and approving the final manuscript. 
We report on the result of `r num_proj` rescue projects that opted to be part of the paper and completed data collection.

A spreadsheet of projects, individual project write-ups (both first replications and rescues), individual project data and analyses for rescue projects, and the analytic code for this paper are all available at https://osf.io/cyk5w/.

@@

The experiments that were re-replicated were chosen from studies that failed to replicate in @boyce2023.
We created an initial list of `r nrow(filtered)` rescue-eligible studies that 1) had received a subjective replication success score of 0, .25, or .5 (on a 0--1 scale) in @boyce2023, 2) had a Github repository available (Github repositories were used in class starting in academic year 2015-2016), and 3) where the original experiment had 200 or fewer participants (to ensure we could afford to match or increase the sample size based on the class budget).
We then contacted the replication project authors for permission to share their report and repository with a new student and include it as a supplement on a resulting paper.
This left `r length(approved)` options for the students to choose from.
`r length(chosen)` students chose to do rescue projects, and
`r length(chosen)-num_proj` students did not finish their project or did not indicate interest in being part of this paper, leaving a final sample of `r num_proj` rescue projects.

@@

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions TODO LINK was given to students.
In general, we encouraged students to add manipulation checks as appropriate, and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of @paxton2012 switched from using the cognitive reflection task (CRT) which many online participants are over-familiar with for the newer CRT-2 which is less overused but attempts to measure the same construct [@thomson2016].
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
@tarampi2016 had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map. The replication ported this experiment online and had participants click a drop down menu to select left or right for each turn. The rescue removed the unwieldy drop down menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

@@

We followed @boyce2023 in the properties of the studies we measured and how we quantified replication success. 
Each project was rated on the basis of subjective replication success (on a 0--1 scale) by both MCF and one of VB and BP.
We thought about replication in terms of general match in key study results (direction, magnitude, and significance), rather than focusing on any singular numeric result or significance cut-off.
Interrater reliability was `r irr |> round(3)`; disagreements were minor (at no point greater than .25) and were resolved through discussion.

@@

We measured statistical consistency using $p$-original, the $p$-value on the null hypothesis that one study's population effect comes from the same distribution of population effects as another study or studies [@mathurNewStatisticalMetrics2020a] If we assume no heterogeneity between studies, then this is the same as asking how consistent one study's estimate of the population-level effect is with another study's estimate of the population-level effect.
However, meta-analyses of closely related studies, such as those from different sites of multi-site replications, indicate that there is often heterogeneity between studies.
$p$-original allows for such heterogeneity, as governed by a parameter $\tau$.
We cannot accurately estimate the level of heterogeneity for an effect meta-analytically as we have only a small number of replications per study.
Instead, we imputed a heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson-collentine2020 in prior multi-site replications in psychology.

@@

For the original study and each replication, we recorded the number of participants post-exclusions.
For studies where some extra conditions were dropped in some replications, we counted only the participants in the key conditions all replications had for comparability.
For instance, if an original study compared between two critical conditions but also had a baseline control, we did not count the participants in the baseline condition if a replication did not include this condition.
We also recorded whether each study was conducted online using a crowd-sourced platform or not.

@@

Our primary question of interest was how many of the `r num_proj` rescue projects succeeded at replicating the results in the original study.
Across the `r nrow(subj)` rescue projects, `r nrow(success)` mostly or fully replicated the original results according to the subjective replication ratings.
`r pluck(tallied,2,1)` got a rating of 0, `r pluck(tallied, 2,2)` got a rating of .75, and `r pluck(tallied, 2,3)` got a rating of 1.
Thus, a first pass answer to the question "how often can a failed replication be salvaged?" is `r round(nrow(success)/nrow(subj)*100)`% (bootstrapped 95% CI: `r round(bootstrap[2]*100)`% - `r round(bootstrap[3]*100)`%) of the time.

@@

smd_scale_small <- ggplot(for_plotting_d |> anti_join(big),aes(x=reorder(name_pretty, point, FUN=max),y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
  scale_shape_manual(values=shape)+
  theme(legend.position = "bottom", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank(),
        legend.title=element_blank(),
        legend.text=element_text(size=10)
        )+
  labs(y="Standarized mean difference", x="")

level_order = c("Yeshurun & Levy 2003",
                "Ngo et al. 2019", 
                 "Krauss & Wang 2003",
                 "Jara-Ettinger et al 2022")
smd_scale_big <- ggplot(for_plotting_d |> inner_join(big) |> 
                          mutate(name_pretty=factor(name_pretty, levels=level_order)), aes(x=name_pretty,y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
    scale_shape_manual(values=shape)+
  theme(legend.position = "none", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standarized mean difference", x="", )

@@

As a complement to subjective replication ratings, we also statistically compared the effect sizes of the rescue, first replication, and original study on one preregistered key measure per study [again following @boyce2023].
Though we did not conduct a systematic review of the literature to identify all possible replications, we also consider the effect sizes obtained in additional replications when we were aware of them (either from other class projects, or external replications in the literature).

@@

all |> ungroup() |>  left_join(names) |> select(Paper=name_pretty, `All reps`=orig_v_other, "Rescue" = orig_v_rescue, "Non-rescue"=orig_v_not_rescue, "Other reps"=rescue_v_reps) |> knitr::kable(format="latex", booktabs=T, linesep="", caption="$P$-original values between different sets of experiments. The primary analysis is between the original result and the meta-analytic aggregation of all replications. All $p$-original values assume an imputed heterogeneity value of $\\tau$=.21.", digits=3) |> 
  add_header_above(c(" "=1, "Original and"=3, "Rescue and"=1)) |> add_header_above(c(" "=1, "p-original comparing between"=4))

@@

Scientists' intuitions about whether a replication is successful and whether an effect provides support for a hypothesis (including heuristic cutoffs like p\<.05) do not always align with measures of statistical consistency [@patil2016].
For instance, two studies may both find that condition 1 results in a significantly higher outcome measure than condition 2, but the effect magnitudes may be sufficiently different that they are statistically unlikely to have come from the same population.
On the other hand, one study may find a statistically significant, but imprecisely estimated effect in a small sample, and a second study may find a near-zero (null) effect, but the effect estimates from the two studies may be statistically compatible, despite one supporting a hypothesized difference and the other not.

As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found).
The value of $p$-original represents how likely an effect size equal to or more extreme than the original effect size is to occur if it comes from a distribution of population effect sizes defined by the mean and standard error of the replications (combined meta-analytically) with a between-population standard deviation of $\tau=.21$ [imputed from @olsson-collentine2020].
That is, $p$-original is a measure of the statistical consistency of the original result with the totality of the replications, assuming a given level of heterogeneity of effects between studies.
Smaller $p$-original values indicate more inconsistency.

@@

As secondary measures, we calculated the $p$-original values between a) the original and the rescue, b) the original and non-rescue replications, and c) the rescue and other replications.
For a) the original versus the rescue, the median value of $p$-original was `r rescue_summ$median |> round(2)` [IQR: `r rescue_summ$low |> round(2)` - `r rescue_summ$high|> round(2)`], and `r round(rescue_summ$less_than*100)`% of the $p$-original values were less than .05.
For b) the original versus all the non-rescue replications, the median value of $p$-original was `r rep_summ$median |> round(2)` [IQR: `r rep_summ$low |> round(2)` - `r rep_summ$high |> round(2)`], and `r round(rep_summ$less_than*100)`% of the $p$-original values were less than .05.
For c) the rescue versus the other replications, the median value of $p$-original was `r rescue_rep_summ$median |> round(2)` [IQR: `r rescue_rep_summ$low |> round(2)` - `r rescue_rep_summ$high |> round(2)`], and `r round(rescue_rep_summ$less_than*100)`% of the $p$-original values were less than .05.

@@

Overall, allowing for a heterogeneity level of $\tau$=.21 SMD, a number of original effects are not statistically consistent with rescue and replication effects.
The pattern of inconsistency does not align with which studies were rated as having replicated.
In all cases, the point estimate of the re-replication is smaller than, or in the opposite direction of, the original effect on the key measure of interest.

@@

Given the mix of successful and unsuccessful rescue projects, we discuss a few projects where we have speculations about why they turned out the way they did.

The rescue of @krauss2003 was successful despite an unsuccessful replication.
This study looked at the influence of guided thinking on whether or not people gave correct justifications (drawn or written) for their answer on the Monty Hall problem.
The original paper reported correct justification from 2/67 (`r round(2/67*100)`%) in the control condition and 13/34 (`r round(13/34*100)`%) in the guided thinking condition.
The first replication struggled to recruit participants who were naive to the problem (an exclusion criterion), and many participants give very short text responses in the provided text box (only textual responses were allowed).
The replication found 0/8 correct justifications in the control and 0/11 in the guided thinking condition.
While we can't know for sure what caused the non-replication, there were clear problems observable from the small final sample and low-quality responses.
The rescue targeted these issues by adding a pre-screening that ensured participants were not familiar with the Monty Hall problem, switched the name of the problem (to reduce googling for answers), and had participants upload drawings for their justifications.
Collectively, we believe these changes brought the rescue closer to the intent of the original.
The rescue had 1/40 (`r round(1/40*100)`%) correct justifications in the control condition and 6/35 (`r round(6/35*100)`%) in the guided thinking condition.
This effect was significant, though smaller than the effect in the original.

@@

@child2018 studied reading times for passages of text with different emotional valences written from either a third-person or second-person perspective.
They found a significant interaction between perspective and valence: participants read positively-valenced passages faster than negatively-valenced passages, but only when they were written from a second-person perspective.
The first replication used a similar sample size to the original study (40) and found an effect that trended in the same direction, but was substantially smaller.
The rescuer identified several possible reasons for this failure, foremost among them that the replication was conducted online while the original was conducted in person.
It is reasonable to expect online studies of reading time to be noisier than in-person studies because the environments participants completed the study in varied much more online.
The rescue attempt was still online, but used a much larger sample size to account for noisier data.
Still, the rescue found a non-significant effect that trended in the opposite direction from the original and first replication.
Even though the first replication trended in the same direction and there was an identifiable reason for the smaller effect size, increasing sample size did not rescue the effect.

@@

We opened  with a question about what an early-career psychology researcher should do given a failed replication: should they try again or give up and move on?
In our sample, the odds of a re-replication working are low [consistent with @ebersole2020].
Especially if there is not a clear, identifiable candidate cause for replication failure, it may be more efficient to select another result to build on.

@@

We thank the students from previous years of PSYCH 251 who graciously agreed to let us re-replicate their projects and shared their work with us. We thank the other members of the PSYCH 251 teaching team and students for feedback on projects. Data collection was supported by funding from the Stanford Departments of Psychology and Philosophy, Graduate School of Education, and the Symbolic Systems program.

@@

VB and MCF conceptualized the study. BP contributed project administration, VB contributed formal analysis, and MCF contributed supervision.

Authors other than VB, BP, and MCF each designed, conducted, analyzed, and wrote up one of the rescue projects -- these authors contributed to investigation, formal analysis, and visualization.

VB wrote the first draft, BP and MCF contributed editing, and all authors reviewed the final manuscript.
